---
title: "Immigrant map of the United States"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
\
\
\


 **We look at the top countries of origin for foreign-born individuals in the United States. We use data from the American Community Survey (ACS) published by the Census Bureau.** 
\
\


```{r libraries, message=FALSE, include=FALSE}
# Load libraries:
library(dplyr)
library(data.table)
library(plotly)
```


```{r, message=FALSE, include=FALSE}
# Load data and select variables:
selected.variables = c("NATIVITY","ST","POBP", "PWGTP")
hus.a.data = fread("./data/ss14pusa.csv", select = selected.variables)
hus.b.data = fread("./data/ss14pusb.csv", select = selected.variables)
db = rbind(hus.a.data, hus.b.data)
rm(hus.a.data, hus.b.data)

# save(db, file = "./data/db.RData")
# load("../data/db.RData")
```



```{r, message=FALSE, include=FALSE}
# Get country names and state abbreviations:

countries = fread("./data/countrynames.csv")
db = db %>% 
    left_join(., countries, by = c("POBP" = "code")) %>%
    rename(COB_name = name)

states = fread("./data/statenames.csv")
db = db %>% 
    left_join(., states, by = c("ST" = "code")) %>% 
    rename(ST_abbr = abbr) %>% 
    select(-c(name))

statenames = states$name
names(statenames) = states$abbr

# remove DC and Puerto Rico rows:
db = filter(db, !ST_abbr %in% c('DC','PR'))
statenames = statenames[!statenames %in% c('District of Columbia','Puerto Rico')]
```


```{r, message=FALSE, include=FALSE}
# Find top countries of immigration:
top_n = 20 # the top how many countries of immigration
top_countries = db %>% 
                filter(NATIVITY==2) %>%
                group_by(COB_name) %>%
                summarise(sum(PWGTP)) %>%
                arrange(desc(`sum(PWGTP)`)) %>%
                head(top_n)
```


### Top `r top_n` countries of origin for immigrants in the United States:

| Rank |   |Country | Immigrant Population|
|------|---|--------|---------------------|
`r paste('|', 1:top_n, '|', paste('![](./data/Flags/24/',top_countries[[1]],'.png)', sep=''), '|', top_countries[[1]], '|', round(top_countries[[2]]/1e6, 3), 'M', '|', '\n', collapse='')`
###### (Icons: <http://www.icondrawer.com>)


```{r, message=FALSE, include=FALSE}
# Estimate percent immigrant in state poulation:

pop_total = db %>% group_by(ST_abbr) %>% summarise(sum(PWGTP)) %>% rename(tot = `sum(PWGTP)`)
pct_imm_by_st_df = db %>% 
                  filter(NATIVITY==2) %>% 
                  group_by(ST_abbr) %>% 
                  summarise(sum(PWGTP)) %>% 
                  mutate(percent_imm = round(100*`sum(PWGTP)`/pop_total$tot, 1)) %>%
                  select(-`sum(PWGTP)`)

pct_imm_by_st = pct_imm_by_st_df$percent_imm
names(pct_imm_by_st) = pct_imm_by_st_df$ST_abbr
```



```{r, message=FALSE, include=FALSE}
# For each US state, get top countries of immigration:
top_n_st = 5 # the top how many countries of immigration to display
top_countries = list()
for(s in unique(db$ST_abbr)){
  state_rows = filter(db, ST_abbr==s & NATIVITY==2) # select only immigrants in state s
  top_countries[[s]] = state_rows %>%
                      group_by(COB_name) %>%
                      summarise(sum(PWGTP)) %>%
                      mutate(pct = round(100* `sum(PWGTP)` /sum(state_rows$PWGTP), 1)) %>%
                      select(-`sum(PWGTP)`) %>%
                      arrange(desc(pct)) %>%
                      head(top_n_st)
}
```
\
\


### Geographic distribution of immigrant groups:
\
\
```{r, message=FALSE, echo=FALSE}
# Display map:

nstates = length(top_countries)
hover = character(nstates)
percent_imm = numeric(nstates)

# prepare list for display of top countries of immigration:
for(i in names(statenames)){
  hover[i] = paste('<b>', statenames[i], ' ', pct_imm_by_st[i], '%</b>, of which:<br>',
                    paste('  ',top_countries[[i]]$COB_name, ' ', top_countries[[i]]$pct, '% <br>', collapse=""),
                    collapse="")
  percent_imm[i] = pct_imm_by_st[i]
}

# combine all in data frame for plotting
state_imm = data.frame(code = names(statenames), hover = hover, percent_imm = percent_imm, stringsAsFactors=FALSE)

# give state boundaries a white border
l = list(color = toRGB("white"), width = 2)

# specify some map projection/options
g = list(scope = 'usa', projection = list(type = 'albers usa'), showlakes = TRUE, lakecolor = toRGB('white'))

plot_ly(state_imm, type = 'choropleth', z = percent_imm, 
        locationmode = 'USA-states', locations = code, 
        hoverinfo = 'text', text = hover, 
        color = percent_imm, colors = 'Oranges',
        marker = list(line = l), 
        colorbar = list(len = 0.75, thickness = 20, title = "Percent foreign-born")) %>%
  
  layout(geo = g, 
         title = 'Percent foreign-born per state and top countries of origin <br> (hover cursor over map for breakdown)', 
         autosize = FALSE,
         width = 900,
         margin = list(t=25, l=0, r=0, d=0))

```

